---
title: "Sierra Acoustics Data Exploration"
output: html_document
date: "2024-09-26"
---

# Package Loading
```{r message = FALSE}
library(here)
library(dplyr)

```

# Thresholds file exploration:

## Variable types
1. Species {character} = Species Common Name
2. keep {numeric} = Binary indicator variable
3. BirdNET_code {numeric} = Species' code names from BirdNET
4. intercept.c {numeric}
5. beta.c {numeric}
6. intercept.r {numeric}
7. beta.r {numeric}
8. cutoff85.r {numeric}
9. cutoff90.r {numeric}
10. cutoff95.r {numeric}
11. cutoff975.r {numeric}
12. cutoff99.r {numeric}
13. cutoff90.r_conf {numeric}
14. cutoff95.r_conf {numeric}
15. cutoff99.r_conf {numeric}

"r" indicates raw/logit scale values and "c|conf" indicates in the probability scale for confidence scores (confidence is not a probability itself). Cutoffs are the value where all data below this value is considered a non-detection (in an occupancy modelling framework this is okay given the structure accounts for false negatives). All values greater than the cut-off is left as the score. False positives are what we are trying to minimize since they are not easily integrated into an occupancy modelling framework.

```{r}
## Read thresholds in
thresh <- readr::read_csv(here("Data/Thresholds_2021_20230309.csv"))
glimpse(thresh)

## DF Dimensions
dim(thresh)

## How many species
length(unique(thresh$species))
print(thresh$species)

## Names of the columns
colnames(thresh)

## Breakdown for "keep"
table(thresh$keep)

## Overall summarization
summary(thresh)

```

## ARU and Species Detection Data
The following data files correspond to Step3 in the BirdNET

```{r}
## Load in an example species file
atfl <- readr::read_csv(here("./Data/Detections_By_Species/2021/Ash-throated_Flycatcher_Gt0.563_2021_max_score_summary.csv"))

## Structure of the dataframe
str(atfl)

## Replace the "." notation with a "0"
atfl[atfl == "."] <- "0"

## Make all characters numeric
atfl <- atfl |> 
  mutate(across(2:last_col(), ~as.numeric(.))) |> #characters to numeric 
  mutate(across(2:last_col(), ~ifelse(. > 0, 1, .))) #make binary

## How many detections of Ash-throated Flycatcher across Sierra
paste("Ash-throated Flycatcher detections:", sum(atfl[2:ncol(atfl)]))

## Total Survey records
(ncol(atfl)-1) * nrow(atfl)

## % Detections
atfl_pct_det <- sum(atfl[2:ncol(atfl)]) / ((ncol(atfl)-1) * nrow(atfl)) * 100

## Percentage of surveys with ATFL
paste("Percent of surveys with ATFL confirmed:", round(atfl_pct_det, digits = 2), "%")

```

## Linking ARU data to spatial data
We now have the ARU data formatted into an ARU x Date matrix with cells corresponding to the BirdNET confidence scores (thresholded to the maximum detection score per day >= the estimated BirdNET thresholds). The cells now represent naive occupancy for the species of interest. Next, we will want to assign the ARUs a spatial location using spatial references for each ARU.

WIP: Checking with Jay to get access to spatial data.




