---
title: "Sierra Acoustics Data Exploration"
output: html_document
date: "2024-09-26"
---

# Package Loading
```{r message = FALSE}
library(here)
library(dplyr)

```

# Thresholds file exploration:

## Variable types
1. Species {character} = Species Common Name
2. keep {numeric} = Binary indicator variable
3. BirdNET_code {numeric} = Species' code names from BirdNET
4. intercept.c {numeric}
5. beta.c {numeric}
6. intercept.r {numeric}
7. beta.r {numeric}
8. cutoff85.r {numeric}
9. cutoff90.r {numeric}
10. cutoff95.r {numeric}
11. cutoff975.r {numeric}
12. cutoff99.r {numeric}
13. cutoff90.r_conf {numeric}
14. cutoff95.r_conf {numeric}
15. cutoff99.r_conf {numeric}

"r" indicates raw/logit scale values and "c|conf" indicates in the probability scale for confidence scores (confidence is not a probability itself). Cutoffs are the value where all data below this value is considered a non-detection (in an occupancy modelling framework this is okay given the structure accounts for false negatives). All values greater than the cut-off is left as the score. False positives are what we are trying to minimize since they are not easily integrated into an occupancy modelling framework.

```{r}
## Read thresholds in
thresh <- readr::read_csv(here("Data/Thresholds_2021_20230309.csv"))
glimpse(thresh)

## DF Dimensions
dim(thresh)

## How many species
length(unique(thresh$species))
print(thresh$species)

## Names of the columns
colnames(thresh)

## Breakdown for "keep"
table(thresh$keep)

## Overall summarization
summary(thresh)

```

