---
title: "Sierra Acoustics Data Exploration"
output: html_document
date: "2024-09-26"
---

# Package Loading
```{r message = FALSE}
library(here)
library(dplyr)
library(purrr)

```

# Thresholds file exploration:

## Variable types
1. Species {character} = Species Common Name
2. keep {numeric} = Binary indicator variable
3. BirdNET_code {numeric} = Species' code names from BirdNET
4. intercept.c {numeric}
5. beta.c {numeric}
6. intercept.r {numeric}
7. beta.r {numeric}
8. cutoff85.r {numeric}
9. cutoff90.r {numeric}
10. cutoff95.r {numeric}
11. cutoff975.r {numeric}
12. cutoff99.r {numeric}
13. cutoff90.r_conf {numeric}
14. cutoff95.r_conf {numeric}
15. cutoff99.r_conf {numeric}

"r" indicates raw/logit scale values and "c|conf" indicates in the probability scale for confidence scores (confidence is not a probability itself). Cutoffs are the value where all data below this value is considered a non-detection (in an occupancy modelling framework this is okay given the structure accounts for false negatives). All values greater than the cut-off is left as the score. False positives are what we are trying to minimize since they are not easily integrated into an occupancy modelling framework.

```{r}
## Read thresholds in
thresh <- readr::read_csv(here("Data/Thresholds_2021_20230309.csv"))
glimpse(thresh)

## DF Dimensions
dim(thresh)

## How many species
length(unique(thresh$species))
print(thresh$species)

## Names of the columns
colnames(thresh)

## Breakdown for "keep"
table(thresh$keep)

## Overall summarization
summary(thresh)

```

## ARU and Species Detection Data
The following data files correspond to Step3 in the BirdNET.
The purpose the following block should be to load the Step3 data and create 
the detection history across the secondary sampling periods. We can write a small
function to accomplish this based on some modifications we want.

```{r}
## Load in an example species file
atfl <- readr::read_csv(here("./Data/Detections_By_Species/2021/Ash-throated_Flycatcher_Gt0.563_2021_max_score_summary.csv"))

## Structure of the dataframe
str(atfl)

## Replace the "." notation with a "0"
atfl[atfl == "."] <- "0"

## Make all characters numeric
atfl <- atfl |> 
  mutate(across(2:last_col(), ~as.numeric(.))) #characters to numeric 

## What is the max value in the df
min.atfl <- atfl[, 2:ncol(atfl)]
min.val <- min(min.atfl[min.atfl > 0])

## Minimum for the threshold
th <- thresh[thresh$species == "Ash-throated Flycatcher",]$cutoff90.r_conf

## Make the detections binary
atfl <- atfl |> 
  mutate(across(2:last_col(), ~ifelse(. > 0, 1, 0))) #make binary
  
## What is the total number of detections at each site
siteDet <- rowSums(atfl[,2:ncol(atfl)])
DetSites <- which(siteDet >= 2)

## How many detections of Ash-throated Flycatcher across Sierra
paste("Ash-throated Flycatcher detections:", sum(atfl[2:ncol(atfl)]))

## Total Survey records
(ncol(atfl)-1) * nrow(atfl)

## % Detections
atfl_pct_det <- sum(atfl[2:ncol(atfl)]) / ((ncol(atfl)-1) * nrow(atfl)) * 100

## Percentage of surveys with ATFL
paste("Percent of surveys with ATFL confirmed:", round(atfl_pct_det, digits = 2), "%")

## Add function to format the species data according to thresholds and other parameters of
## interest.
ac_det_filter <- function(d,
                          d_thresh,
                          thresh_scale = "Conf",
                          thresh_cut = "99",
                          time_format = "ymd",
                          species,
                          no_dets = 1,
                          binary = T) {
  ## Retrieve the threshold from the species of interest
  sp.thresh <- d_thresh[d_thresh$species == species, ]
  
  ## Check the threshold type
  if (thresh_scale == "Conf") {
    tmp.cols <- which(stringr::str_detect(colnames(sp.thresh), pattern = "conf"))
    tmp.cols <- sp.thresh[, tmp.cols]
    if (all(stringr::str_detect(colnames(tmp.cols), thresh_cut))) {
      print("Threshold not named in threshold dataframe. Check column names and 'thresh_cut'.")
      tmp.thresh <- NA
    } else {
      tmp.thresh <- pull(tmp.cols[which(stringr::str_detect(colnames(tmp.cols), thresh_cut))])
    }
  } else if (thresh_scale == "Raw") {
    tmp.cols <- which(
      stringr::str_detect(colnames(sp.thresh), pattern = "(cutoff)([0-9]+)(.r$)|(cutoff)([0-9]+)(raw$)")
    )
    tmp.cols <- sp.thresh[, tmp.cols]
    if (all(stringr::str_detect(colnames(tmp.cols), thresh_cut))) {
      print("Threshold not named in threshold dataframe. Check column names and 'thresh_cut'.")
      tmp.thresh <- NA
    } else {
      tmp.thresh <- pull(tmp.cols[which(stringr::str_detect(colnames(tmp.cols), thresh_cut))])
    }
  }
  
  ## Check the acoustic data to make sure the threshold is applied
  ## Retrieve the columns names
  colnames(d)
  
  ## Acoustic data will likely contain date specific capture histories...these are the primary interest
  if(time_format == "ymd"){
    date.cols <- colnames(d)[which(
      stringr::str_detect(colnames(d), pattern = "(\\d{4})[[:punct:]]?(\\d{2})[[:punct:]]?(\\d{2})")
    )]
    other.cols <- colnames(d)[which(
      !stringr::str_detect(colnames(d), pattern = "(\\d{4})[[:punct:]]?(\\d{2})[[:punct:]]?(\\d{2})")
    )] 
  } else if(time_format %in% c("dmy", "mdy")){
    date.cols <- colnames(d)[which(
      stringr::str_detect(colnames(d), pattern = "(\\d{2})[[:punct:]]?(\\d{2})[[:punct:]]?(\\d{4})")
    )]
    other.cols <- colnames(d)[which(
      !stringr::str_detect(colnames(d), pattern = "(\\d{2})[[:punct:]]?(\\d{2})[[:punct:]]?(\\d{4})")
    )] 
  } else if(is.na(time_format)){
    stop("Time format not specified.")
  }

  ## We want to operate on the data within the date columns
  ## Find the minimum value for the df
  min.val <- suppressWarnings(min(apply(as.matrix(d[,date.cols]), 1, function(x) {min(x[x>0], na.rm = T)})))
  
  ## If binary
  if(binary) {
    d.thresh <- d |>
      mutate(across(all_of(date.cols), ~ ifelse(. >= tmp.thresh, 1, 0)))
  } else {
    d.thresh <- d |>
      mutate(across(all_of(date.cols), ~ ifelse(. >= tmp.thresh, ., 0)))
  }
  
  ## If the number of detections in the season matters
  if (is.numeric(no_dets)) {
    if (!binary) {
      tmp.detect <- d.thresh |>
        mutate(NDets = rowSums(across(all_of(date.cols)) > 0)) |> #sum logical > 0
        mutate(across(all_of(date.cols), ~ if_else(NDets >= no_dets, ., 0))) |>
        mutate(NDetsNew = rowSums(across(all_of(date.cols)) > 0))
    }
    
    if (binary) {
      tmp.detect <- d.thresh |>
        mutate(NDets = rowSums(across(all_of(date.cols)))) |>
        mutate(across(all_of(date.cols), ~ if_else(NDets >= no_dets, ., 0))) |>
        mutate(NDetsNew = rowSums(across(all_of(date.cols))))
    }
    
    ## Check the columns that changed
    ## If the statement below is true...this means the sum of all the scores below were
    ## less than the number of detections we need at the site over the season and these
    ## rows were made to be entirely nondetections
    all(dplyr::setdiff(tmp.detect$NDets, tmp.detect$NDetsNew) < (no_dets))
    
    ## Drop the helper columns and return d.thresh as the fixed df
    d.thresh <- tmp.detect |>
      dplyr::select(all_of(c(other.cols, date.cols)))
    
    ## Remove superfluous DFs made
    rm(tmp.detect)
  }
  return(d.thresh)
  
} #function

## Test the function
atfl.test99 <- ac_det_filter(d = atfl,
                          d_thresh = thresh,
                          thresh_scale = "Conf",
                          thresh_cut = "99",
                          time_format = "ymd",
                          species = "Ash-throated Flycatcher",
                          no_dets = 2,
                          binary = T)

atfl.test90 <- ac_det_filter(d = atfl,
                          d_thresh = thresh,
                          thresh_scale = "Conf",
                          thresh_cut = "90",
                          time_format = "ymd",
                          species = "Ash-throated Flycatcher",
                          no_dets = 2,
                          binary = T)

atfl.test90.4dets <- ac_det_filter(d = atfl,
                          d_thresh = thresh,
                          thresh_scale = "Conf",
                          thresh_cut = "90",
                          time_format = "ymd",
                          species = "Ash-throated Flycatcher",
                          no_dets = 4,
                          binary = T)


## # of detections
paste("Ash-throated Flycatcher detections:", sum(atfl.test99[2:ncol(atfl.test99)]))
paste("Ash-throated Flycatcher detections:", sum(atfl.test90[2:ncol(atfl.test90)]))
paste("Ash-throated Flycatcher detections:", sum(atfl.test90.4dets[2:ncol(atfl.test90.4dets)]))

## Change in detections
paste("Percent Change in Ash-throated Flycatcher detections from filtering:", ((sum(atfl.test90[2:ncol(atfl.test90)]) - sum(atfl.test99[2:ncol(atfl.test99)])) / sum(atfl.test99[2:ncol(atfl.test99)]) * 100))

## % Detections
atfl_pct_det <- sum(atfl[2:ncol(atfl)]) / ((ncol(atfl)-1) * nrow(atfl)) * 100

## Percentage of surveys with ATFL
paste("Percent of surveys with ATFL confirmed:", round(atfl_pct_det, digits = 2), "%")

```
## Linking ARU data to spatial data
We now have the ARU data formatted into an ARU x Date matrix with cells corresponding to the BirdNET confidence scores (thresholded to the maximum detection score per day >= the estimated BirdNET thresholds). The cells now represent naive occupancy for the species of interest. Next, we will want to assign the ARUs a spatial location using spatial references for each ARU.

Spatial data can be accessed via the `CAbioacoustics` package under the table "acoustic_field_visits".

```{r}
library(CAbioacoustics)
library(sf)

## Connect to the database
## Need to be connected to VPN for this to connect
cb_connect_db()

## Set some values for filters
syear <- 2021
eyear <- 2023

## SF needs to be loaded to execute
ownership <- c('usfs')
cell_list <- cb_cells_by_ownership(ownership)

## Study type
study <- c('Sierra Monitoring')

## Query and pull the table we want
deployments_df <- 
  conn |> 
  dplyr::tbl('acoustic_field_visits') |> 
  dplyr::filter(
    is_invalid == 0,
    cell_id %in% cell_list,
    study_type %in% study,
    survey_year >= syear,
    survey_year <= eyear
  ) |> 
  collect() |> 
  dplyr::select(deployment_name, survey_year, matches("utm"))

## Disconnect from the DB
cb_disconnect_db()
```

### Spatial Plotting for ARUs
`deployments_df` now contains spatial information on the ARUs during the 2021-2023 field deployement. We need to make this into a simple features (SF) object for downstream use.
```{r}

## SF creation
deployments_sf <- 
  deployments_df |> 
  group_split(utm_zone) |> 
  map_dfr(cb_make_aru_sf)

## Plot with MV
## hexes
hexes_map <-
  cb_get_spatial('sierra_hexes') |>
  mapview(layer.name = 'Hexes')

## deployments
deployments_map <-
  deployments_sf |>
  mutate(survey_year = as.factor(survey_year)) |>
  mapview(zcol = 'survey_year', layer.name = 'Survey year')

## combine
hexes_map + deployments_map
```

## Link Species Detections with ARU locations - From Jay's DB
```{r}
glimpse(atfl)
head(atfl$Cell_Unit)
length(unique(atfl$Cell_Unit))


glimpse(deployments_sf)

## Take only the 2021 spatial data
dep21 <- deployments_sf |> 
  filter(survey_year == 2021) |> 
  mutate(Cell_Unit = stringr::str_remove(deployment_name, "G[0-9]+_V[0-9]+_")) |> 
  st_transform(crs = 4326) |> 
  dplyr::mutate(lon = sf::st_coordinates(geometry)[,2],
                lat = sf::st_coordinates(geometry)[,1]) |> 
  st_drop_geometry()

## Check the number of unit cell_unit combinations
## Check duplications in the cell_units
## Duplicates arise via different deployment_ids but not cell_unit ids
## V1 vs V2 in the deployment_id
## Step3 data only contains the cell_unit ID
## Deployments from owl DB have 1548 unique deployment_ids
## ATFL example species from 2021 has 1652 individual ARU deployments
nrow(dep21)
length(unique(dep21$deployment_name))
length(unique(dep21$Cell_Unit))
dup_unit <- dep21[which(duplicated(dep21$Cell_Unit)),]$Cell_Unit
dup21 <- dep21[dep21$Cell_Unit %in% dup_unit, ]


## Match the ATFL data with the 2021 survey data
atfl.jay <- atfl |>
   mutate(Cell_Unit = ifelse(
    stringr::str_detect(string = Cell_Unit, pattern = "C[0-9]{3}"),
    gsub(pattern = "(C)([0-9]{3})(_U[0-9]+)$", replacement = "\\10\\2\\3", x = Cell_Unit),
    Cell_Unit
  )) |> 
  left_join(dep21)

missing <- dep21[!dep21$Cell_Unit %in% atfl.jay$Cell_Unit,]

atfl.sf.ok <- atfl.sf[which(!is.na(atfl.jay$survey_year)),]
atfl.sf.ok <- atfl.sf.ok |> 
  dplyr::select(Cell_Unit, deployment_name, survey_year, lon, lat, everything()) |> 
  rowwise() |> 
  mutate(sum = sum(c_across("2021-03-01":"2021-08-30"))) |> 
  ungroup() |> 
  dplyr::select(Cell_Unit, deployment_name, survey_year, lon, lat, sum) |> 
  st_as_sf(coords = c("lon", "lat"), crs = 4326)
  
ggplot(data = atfl.sf.ok) + 
  geom_sf(aes(color = sum))



```

## Data file with locations from Connor
It seems like there are some inconsistencies with the leading zero in the cell ID name. My guess is that importing the files into excel dropped the leading zeros since all of the issues are occurring with cells that have 3 numbers when they should have 4. Add in a leading zero if the string only has 3 numbers after the "C" and let's see what happens.

```{r}
aru_meta <- readr::read_csv(here("Data/ARU_120m.csv"))
names(aru_meta)
glimpse(aru_meta)
aru_meta$Cell_Unit <- paste0(aru_meta$cell_id, "_", aru_meta$unit_numbe)

## Take a couple of things for testing
aru_meta <- aru_meta |>
  dplyr::select(Cell_Unit, survey_yea, X, Y, cc_cfo_mn)

## Attempt to merge this with the ATFL dataset
atfl_meta <- d.thresh |> 
  left_join(aru_meta)

## Fixing the leading zero issue
atfl.zeros <- d.thresh |>
  mutate(Cell_Unit = ifelse(
    stringr::str_detect(string = Cell_Unit, pattern = "C[0-9]{3}"),
    gsub(pattern = "(C)([0-9]{3})(_U[0-9]+)$", replacement = "\\10\\2\\3", x = Cell_Unit),
    Cell_Unit
  ))

## Attempt to rejoin now
atfl_meta <- atfl.zeros |> 
  left_join(aru_meta)

## Issue fixed
sum(is.na(atfl_meta$survey_yea))

## Make the ATFL data an SF object
class(atfl_meta)
atfl.sf <- atfl_meta |> 
  st_as_sf(coords = c("X", "Y"), crs = 4326)

atfl.sf |>
  mapview(zcol = '2021-06-15', layer.name = 'ATFL Detections')

```
